>

### 2020-03-16 manuscript/manuscript.Rmd
Update formulae
Reviewer 2 asked that we be more consistent with our use of matrix and vector annotation. This commit places an explicit boldsymbol around all the variables representing vectors in all equations. Closes #6

A third accessibility mechanism is a utility-based specification, termed as such
for being derived from random utility choice theory. Consider that an individual
living at point $i$ is choosing a park for a recreation activity. If we apply
the multinomial logit model [@McFadden1974], the expected consumer surplus
enjoyed by this individual can be shown to be
\begin{equation}\label{eq:utility}
A_i = \ln\left({\sum_{j = 1}^J\exp(V_{ij})}\right) + C
\end{equation}
where parks are differentiated from each other by their relative measurable
*utilities*, $V_{ij}$. $C$ is an unknown constant, but the difference in consumer
surplus between two points $i$ and $k$ can be quantitatively compared as 
$A_i - A_k$ [@Bruce1977]. In principle, $V$ may include any measurable attributes of either
the choice maker or the park, and is typically represented as a linear-in-parameters
function of destination attributes $X_j$ and the travel cost $d_{ij}$,
\begin{equation}\label{eq:utilityV}
V_{ij} = X_j\beta + \beta_d * d_{ij}
\end{equation}
The coefficients $\vect{\beta}$ are frequently estimated from household
surveys, though in the absence of a survey we may assert reasonable values.

Note that the gravity formulation is itself a special case of the utility-based
specification where $\exp(X_j\beta + \beta_d d_{ij}) = X_j f(d_{ij})$
[@Daly1982]. The distinction between gravity and utility-based specification is
meaningful, however. Primarily, it becomes possible to construct accessibility
statistics based on revealed behavioral preferences rather than calibrated or
asserted values [@Handy1997]. @Kaczynski2016 present what is effectively a 
utility-based accessibility score they call "ParkIndex", with the explicit
motivation of developing a uniformly applicable park accessibility statistic,
though they limit this statistic as a way to add heterogeneity within an
isochrone analysis.

@@

We then calculate the utility-based accessibility of each tract $A_i$ as defined
in Equation \ref{eq:utility} under each utility specification. Recall that the 
total value of the accessibility is relative to an unknown scalar $C$; for this 
reason we standardize the accessibility values for all tracts within each
utility specification, 
\begin{equation}\label{eq:v_logged}
A_i' = \frac{A_i - \bar{A}}{sd(A)}
\end{equation}

@@

We predict either the physical activity or the obesity rate ($y$) in a census
tract as a linear function of the tract's sociodemographic characteristics $X$
and accessibility to parks $A$,
\begin{equation}\label{eq:themodel}
 y = X\beta + \beta_{a}A + \varepsilon
\end{equation}

As the observations are related to each other spatially, it is likely that this
process involves spatial spillovers of one kind or another. A complete treatment of
spatial econometrics is not warranted here; the reader is directed to
@LeSage2009 as well as @LeSage2014.
Suffice it to say that the presence of spatial data generating processes can
negatively affect econometric interpretation in a variety of ways. Each of these
processes rely on spatial autoregression, where elements of some random
variable $x$ are spatially dependent on other elements, 
\begin{equation}\label{eq:autoregression}
 x = \rho W x + \varepsilon
\end{equation}
with the strength of dependence an estimated parameter $\rho$, the structure of 
spatial dependence governed by the asserted matrix $W$ (details of $W$ follow below),
and some random uncorrelated residual $\varepsilon$. A collection of models is
available to represent spatial autodependence in the model residuals, the 
independent variables, and the dependent variable.

The specific spatial model applied is both an econometric choice
--- as improperly specified spatial models can lead to inconsistent estimates of
model parameters and standard errors --- as well as a philosophical choice about
the likely structure of spillovers in the problem at hand. In case of physical
activity and obesity rates, we believe that the outcomes are locally dependent:
That is, a particular individual decides whether or not to participate in
physical activity independently of whether his or her neighbors
participate in physical activity. This does not mean that we think it impossible that
the socioeconomic variables of neighboring tracts do not play a role. Mathematically, we assert that the autodependence
relationship on the dependent variable $y = \rho W y + \ldots$ has 
$\rho = 0$ in the current case.

With this relationship ruled out, it remains a possibility that the 
model residuals are spatially dependent,
\begin{equation}\label{eq:SEM}
 y = X\beta + \beta_a A + u; u = \lambda W u + \varepsilon
\end{equation}
that the outcome is partially dependent on the socioeconomic variables in *neighboring*
tracts,
\begin{equation}\label{eq:SLX}
 y = X\beta + \beta_a A + W X \gamma + \varepsilon
\end{equation}
or a linear combination of both.
\begin{equation}\label{eq:SDEM}
 y = X\beta + \beta_a A + W X \gamma + u; u = \lambda W u + \varepsilon 
\end{equation}
These models are referred as the spatial error model (SEM), the spatial lag of
$X$ model (SLX), and the spatial Durbin error model (SDEM). @Pace2008 suggest
that a Hausman-style test of the estimates of $\beta$ derived from the OLS 
(Equation \ref{eq:themodel}) and SEM (Equation \ref{eq:SEM}) can identify whether
the estimates are consistent. If the $\beta$ estimates are dissimilar, both 
specifications are inappropriate as they contain untreated omitted variables.
Similarly, if the SLX and SDEM estimates are dissimilar it is an indication of 
missing variables in that specification. On the other hand if the estimates of
both pairs of models are similar and the residual correlation parameter $\lambda$
is non-zero, the model accounting for spatially dependent residuals will produce 
proper estimates of the model standard errors.

@@

The coefficients in the SDEM model are of two general types: the direct effect
resulting from a tract's own attributes and the indirect effect ($\gamma$)
resulting from the spatially-weighted attributes of its neighbors. For example, a
percentage point increase in the share of young adults in a tract increases the
expected physical activity rate in a tract by 
`r slx_vs_sdem[[2]]$coefficients["young_adults"]` percentage points, and a
percentage point increase in the share of young adults
*in neighboring tracts* increases the expected physical activity rate by an 
*additional* `r slx_vs_sdem[[2]]$coefficients["lag.young_adults"]` percentage
points. This opportunity to examine spatial spillovers of attributes is a key 
benefit of using a spatial model.

>>>>

A third accessibility mechanism is a utility-based specification, termed as such
for being derived from random utility choice theory. Consider that an individual
living at point $i$ is choosing a park for a recreation activity. If we apply
the multinomial logit model [@McFadden1974], the expected consumer surplus
enjoyed by this individual can be shown to be
\begin{equation}\label{eq:utility}
A_i = \ln\left({\sum_{j = 1}^J\exp(V_{ij})}\right) + C
\end{equation}
where parks are differentiated from each other by their relative measurable
*utilities*, $V_{ij}$. $C$ is an unknown constant, but the difference in consumer
surplus between two points $i$ and $k$ can be quantitatively compared as 
$A_i - A_k$ [@Bruce1977]. In principle, $V$ may include any measurable attributes of either
the choice maker or the park, and is typically represented as a linear-in-parameters
function of destination attributes $X_j$ and the travel cost $d_{ij}$,
\begin{equation}\label{eq:utilityV}
V_{ij} = X_j\boldsymbol{\beta} + \beta_d * d_{ij}
\end{equation}
The coefficients $\boldsymbol{\beta}$ are frequently estimated from household
surveys, though in the absence of a survey we may assert reasonable values.

Note that the gravity formulation is itself a special case of the utility-based
specification where $\exp(X_j\boldsymbol{\beta} + \beta_d d_{ij}) = X_j f(d_{ij})$
[@Daly1982]. The distinction between gravity and utility-based specification is
meaningful, however. Primarily, it becomes possible to construct accessibility
statistics based on revealed behavioral preferences rather than calibrated or
asserted values [@Handy1997]. @Kaczynski2016 present what is effectively a 
utility-based accessibility score they call "ParkIndex", with the explicit
motivation of developing a uniformly applicable park accessibility statistic,
though they limit this statistic as a way to add heterogeneity within an
isochrone analysis.

@@

We then calculate the utility-based accessibility of each tract $A_i$ as defined
in Equation \ref{eq:utility} under each utility specification. Recall that the 
total value of the accessibility is relative to an unknown scalar $C$; for this 
reason we standardize the accessibility values for all tracts within each
utility specification, 
\begin{equation}\label{eq:v_logged}
A_i' = \frac{A_i - \bar{\boldsymbol{A}}}
            {\sigma_{\boldsymbol{A}}}
\end{equation}

@@

We predict either the physical activity or the obesity rate ($y$) in a census
tract as a linear function of the tract's sociodemographic characteristics $X$
and accessibility to parks $A$,
\begin{equation}\label{eq:themodel}
 \boldsymbol{y} = X\boldsymbol{\beta} + \beta_{a}\boldsymbol{A} + \boldsymbol{\varepsilon}
\end{equation}

As the observations are related to each other spatially, it is likely that this
process involves spatial spillovers of one kind or another. A complete treatment of
spatial econometrics is not warranted here; the reader is directed to
@LeSage2009 as well as @LeSage2014.
Suffice it to say that the presence of spatial data generating processes can
negatively affect econometric interpretation in a variety of ways. Each of these
processes rely on spatial autoregression, where elements of some random
variable $x$ are spatially dependent on other elements, 
\begin{equation}\label{eq:autoregression}
 \boldsymbol{x} = \rho W \boldsymbol{x} + \boldsymbol{\varepsilon}
\end{equation}
with the strength of dependence an estimated parameter $\rho$, the structure of 
spatial dependence governed by the asserted matrix $W$ (details of $W$ follow below),
and some random uncorrelated residual $\varepsilon$. A collection of models is
available to represent spatial autodependence in the model residuals, the 
independent variables, and the dependent variable.

The specific spatial model applied is both an econometric choice
--- as improperly specified spatial models can lead to inconsistent estimates of
model parameters and standard errors --- as well as a philosophical choice about
the likely structure of spillovers in the problem at hand. In case of physical
activity and obesity rates, we believe that the outcomes are locally dependent:
That is, a particular individual decides whether or not to participate in
physical activity independently of whether his or her neighbors
participate in physical activity. This does not mean that we think it impossible that
the socioeconomic variables of neighboring tracts do not play a role. Mathematically, we assert that the autodependence
relationship on the dependent variable $\boldsymbol{y} = \rho W \boldsymbol{y} + \ldots$ has 
$\rho = 0$ in the current case.

With this relationship ruled out, it remains a possibility that the 
model residuals are spatially dependent,
\begin{equation}\label{eq:SEM}
 \boldsymbol{y} = X\boldsymbol{\beta} + \beta_a \boldsymbol{A} + \boldsymbol{u}; \boldsymbol{u} = \lambda W \boldsymbol{u} + \boldsymbol{\varepsilon}
\end{equation}
that the outcome is partially dependent on the socioeconomic variables in *neighboring*
tracts,
\begin{equation}\label{eq:SLX}
 \boldsymbol{y} = X\boldsymbol{\beta} + \beta_a \boldsymbol{A} + W X \boldsymbol{\gamma} + \boldsymbol{\varepsilon}
\end{equation}
or a linear combination of both.
\begin{equation}\label{eq:SDEM}
 \boldsymbol{y} = X\boldsymbol{\beta} + \beta_a \boldsymbol{A} + W X \boldsymbol{\gamma} + \boldsymbol{u}; \boldsymbol{u} = \lambda W \boldsymbol{u} + \boldsymbol{\varepsilon}
\end{equation}
These models are referred as the spatial error model (SEM), the spatial lag of
$X$ model (SLX), and the spatial Durbin error model (SDEM). @Pace2008 suggest
that a Hausman-style test of the estimates of $\boldsymbol{\beta}$ derived from the OLS 
(Equation \ref{eq:themodel}) and SEM (Equation \ref{eq:SEM}) can identify whether
the estimates are consistent. If the $\boldsymbol{\beta}$ estimates are dissimilar, both 
specifications are inappropriate as they contain untreated omitted variables.
Similarly, if the SLX and SDEM estimates are dissimilar it is an indication of 
missing variables in that specification. On the other hand if the estimates of
both pairs of models are similar and the residual correlation parameter $\lambda$
is non-zero, the model accounting for spatially dependent residuals will produce 
proper estimates of the model standard errors.

@@

The coefficients in the SDEM model are of two general types: the direct effect
resulting from a tract's own attributes and the indirect effect ($\boldsymbol{\gamma}$)
resulting from the spatially-weighted attributes of its neighbors. For example, a
percentage point increase in the share of young adults in a tract increases the
expected physical activity rate in a tract by 
`r slx_vs_sdem[[2]]$coefficients["young_adults"]` percentage points, and a
percentage point increase in the share of young adults
*in neighboring tracts* increases the expected physical activity rate by an 
*additional* `r slx_vs_sdem[[2]]$coefficients["lag.young_adults"]` percentage
points. This opportunity to examine spatial spillovers of attributes is a key 
benefit of using a spatial model.
